An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network

### 2025-07-16 paper/main.tex
Update main.tex

\subsection*{Automatic Mixed Precision and Gradient Clipping}
As a next experiment, \dots %Sayed, also use the other images?
\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/gradclip_32.png}
	\caption{Inference accuracy using float32-trained weights with AMP and gradient clipping.}
\end{figure}

>>>>

\subsection*{Automatic Mixed Precision and Gradient Clipping}
Automatic Mixed Precision (AMP) accelerates deep learning training by using both float16
and float32 data types reducing memory usage and increasing computational speed. 
However, float16’s limited dynamic range can cause numerical instability,
especially during backpropagation. To address this, Gradient Clipping is used to cap 
the magnitude of gradients—either globally or per element—preventing extreme updates
that could hinder convergence. Together, AMP and gradient clipping enable efficient,
stable training by balancing performance and numerical precision.

First step is using global norm clipping, all gradients are collectively scaled to ensure
their combined norm does not exceed a specified threshold, promoting stable updates
across the network. In second step per-element clipping constrains each individual gradient
component within a fixed range, directly limiting extreme values.
These methods are evaluated under Automatic Mixed Precision (AMP) to compare their impact on
training stability, model accuracy, and computation time across different precision types.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/FP16_32_clipping.png}
	\caption{Affect of Gradient Norm on accuracy, training and testing timings.
                 Accuracy is virtually the same between FP16 and FP32 when gradient clipping is used
		 validating the stability of AMP and gradient scaling.
                 Eval time is nearly identical for both formats, as evaluation involves fewer operations
		 and less intense memory bandwidth usage.}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/gradclip_16.png}
	\caption{Effect of Per-Element Gradient Clipping on FP16.
                 Accuracy vs Training & Eval Time}
        \label{fig:grad16}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.85\textwidth]{figures/gradclip_32.png}
	\caption{Effect of Per-Element Gradient Clipping on FP32.
                 Accuracy vs Training & Eval Time}
        \label{fig:grad32}
\end{figure}

Figure~\ref{fig:grad16} shows accuracy peaked at 97.07% with a clipping value of 0.5, 
while the lowest accuracy (96.55%) occurred at clip=0.01, suggesting that very small
clip values may overly restrict gradient updates and hurt learning.
Training time ranged indicating that moderate clipping supports efficient convergence.
Evaluation time remained stable across all clip values (around 2.01s to 2.16s),
showing that clipping primarily impacts training rather than inference performance. 
Also Figure~\ref{fig:grad32} shows that accuracy was most stable across all clip values
with the highest accuracy at clip=0.5, demonstrating its robustness to gradient clipping
variations. Training time suggesting that larger clip values can speed up convergence
without sacrificing stability.
Evaluation time confirming that clipping had minimal effect on inference time.
